### 2015-04-03 Workshop2015/current-version/incremental.tex
new current version

To answer the needs expressed above, we propose an SLA guided, security-aware continuous data provision and integration approach performed within a multi-cloud service oriented environment  (see
%starting from the processing of the query  to the delivery of the results.
Figure~\ref{fig:arch} )
%shows the general architecture of an SLA guided data integration system. 
%It accesses data services which are data providers deployed in a cloud  that provide agreed SLA's. 
%The whole process is monitored to determine whether a computed SLA is being honoured while a query is evaluated.

@@

Services are deployed in different cloud providers and each service exports an agreed SLA that specifies the economic cost per call, the maximum number of calls that can be done per day, the availability of the service, the average response time when a method is called, the reliability, the privacy of the produced data (whether they can be stored or not), the precision, freshness and provenance of the produced data.

@@

Some of these measures ({\sf cost/call, maxCall/day}) are static and explicitly specified by the service provider. 
In contrast, the other measures should be computed by monitoring the conversations between the service and the applications that contact it.  It is the case for service reputation that will condition the way personal user data will be anonymized (Constraint1).  Some measures will be instantiated depending on the context of  the service invoked within a multi-cloud environment. It is the case for constraint 2 expressed above:   user data access credentials should be calculated according to the clouds hosting different data services.

According to our approach, content producers are modeled as data services with associated \textit{agreed SLAs}. 
In our example of MOOC system presented above, we assume that several producers will supply equivalent content for a given period of time  that will be chosen with respect to the  preferences expressed by a consumer. 
Assume that there are four content providers on English poetry that can be queried individually and two hubs that collect information from other sources like social network groups and hash tags.
Hubs will store content about given topics on  English poetry, available from particular producers (e.g., participants of a given course).
We represent these services by {\sf e$_1$, \dots, e$_4$} and {\sf h$_1$} and {\sf h$_2$}. We suppose that hubs will have the same interface as individual service providers.
We also suppose the existence of two free location services exporting the following interface: {\sf loc(IP) $\rightarrow$ $\langle$ X, Y$\rangle$}, meaning that given an IP address it returns a geographic position expressed as a pair of coordinates. 
All these services can potentially be combined for answering queries.

In our example, non-functional characteristics of the content provision services ({\sf e$_1$, ... e$_4$}) are calculated by services that enforce SLAs  as follows:
\begin{trivlist}\sf\footnotesize
\item[~-~e$_i$:]estim-cost$_i$(contents, req\_size, cost, prov\_size, loc), QoS preferences$_\mathit{user}$.

@@

\color{red}

The first operation permits the user to ask for the budget of a given course (content) and a required minimmum size.
This operation will return the cost of the data, as well as its actual size and location.
{QoS preferences}$_\mathit{user}$ is the user's proposal to the contents provider. 
This SLA will become the {\sf agreedSLA}$_i$, in case of availability of the contents. 
We suppose that availability is characterized by a non-zero value of size of the provided data ({\sf prov\_size}).
The second operation is used to engage in a course. 
The agreed SLAs are those containing the information described above in this example.
%The user will ask for a given derived SLA to which the provider may agree.

Cloud providers also define their SLA contracts that specify the cost per request ({\sf cost/request}), the volume of data that can be exchanged per month ({\sf I/0 volume/month}), the cost of transferring data or applications within the same data center or between data centers ({\sf datatransferCost/region}), the storage space ({\sf storageSpace}). For example some cloud providers enable the customer to choose the zone to install PaaS services and deploy applications (e.g. zone 1 is Europe). If the customer wishes to deploy services in zone 1 but store data in zone 2 the transfer cost will change.

@@

@@

Queries may be expressed as Datalog-like programs or an SQL-like expressions, including spatio-temporal attributes and preferences.
For instance, let us have the query ``\textit{List of English poetry content providers that can provide commented Emily Dickinson poems that are close to my city and that are labelled as experts, where the total cost is not higher to 1 dollar}''. 
This query has two parts: the one that corresponds to the content that will be solved by combining data services and the user preferences. For this query the user is ready to pay a maximum of \$1 as total cost; she requests that content providers be certified as experts (provenance) and that her privacy be saved by adapting the anonymity process with regard to services reputation. The maximum total cost will condition also which kind of privacy preserving algorithms to apply.

The user preferences statement would include their cloud provider contract and quality preferences:
\begin{trivlist}\sf\footnotesize
\item[~-~cloudSLA:]  $\langle$0,05 cents per call, 8 GB I/0 volume/month, free, 1 GB storage$\rangle$. 
\end{trivlist}

We consider a simplified cloud SLA contract inspired in the lowest contract provided by Azure\footnote{Azure is a trademark of Microsoft Corporation.}: {\sf cost of \$0,05 cents per call,  8~GB of I/O volume/month, free data transfer cost within the same region,  01~GB of storage}.

The user is ready to pay a maximum of {\sf \$1 as total query cost}; she requests that only {\sf certified} content providers are listed (provenance); even if their data are not fresh; she requires a contract duration of one week.

\begin{trivlist}\sf\footnotesize
\item[~-~SLA$_\mathit{user}$: ] $\langle$cost $\leq$ \$1, availability, freshness = ``any'', provenance = ``certified'', location = ``close'', duration = 7 days, privacy-preserving=``reputation-based''$\rangle$. 
\end{trivlist}

As usual, SLAs will be represented as conditions over variables that will be used in the query.
In this context, the compliance to the SLAs will drive the query rewriting process.

Given a requirement expressing a query and quality of service preferences: cost, provenance, trust requirement, execution time, data confidentiality needs, \dots, the data integration process for computing a result is done as follows: (i) User preferences (including cloud SLAs according to user subscriptions) and agreed SLAs will be used to produce a derived SLA to be associated to the query. 
These derived SLA will influence the choice of (data) services that will actually be used for building the result; (ii) Query rewriting for computing service compositions that can be used for  building query results. The agreed SLA proposed by the chosen services is added as  conditions in  each re-written   query expression; (iii) Managing the integration process particularly storing partial results (including in  which location data is stored), delivering data (applying security control during integration) and launching and re-launching queries. Intermediate results  are stored as knowledge in order to reduce the overhead of the query evaluation process.

\color{black}
%-----------------------------------------------------------------
\section{Derived SLA}
\label{sec:slaModel}
%-----------------------------------------------------------------

The key and original aspect of   our proposed data integration and provision process is  defined as a vertical mapping of user QoS preferences and agreed SLAs. This  leads to a {\em derived SLA} that guides the evaluation of a query.

A query has associated preferences  expressed as macroscopic constraints (i.e. user preferences statement): execution time, pay / no pay, data reliability, provenance, freshness, privacy-preserving constraints, partial/full results, delivery mode. These constraints are coupled with the profile of the user which is in general stated in her cloud subscription (amount of assigned storage space, number of requests, I/O transfered Mega bytes, etc.).

As said before, we assume that services export SLAs (i.e., agreed SLAs) that define measures   that can be either expressed as constants,  computed (dynamically) by monitoring the execution and conversations associated to services, and hybrid they can be statically stated  but they change at execution time.  A service  agreed SLA is expressed through an  XML document using the specification WSLA (Web service level agreement \footnote{\footnotesize http://www.research.ibm.com/people/a/akeller/\-Data/WSLASpecV1-20030128.pdf}). The service SLA measures  that we consider in our example are: cost, availability, freshness, provenance, location, duration of the engagement, privacy-preserving status. Other measures are associated to the conditions in which the service is called or to the precision and recall of their produced data given a request.

@@

%-----------------------------------------------------------------
%\section{Dealing with the resources consumed by the evaluation process}
%\label{sec:queryProcessOpt}
%-----------------------------------------------------------------
 %Our data provision and integration approach relies on data services deployed on one or different clouds providers and it is delivered as a distributed DaaS on top of the clouds involved.  This DaaS  uses resources from a cloud and this use  is  guided by an economic model (stated in a cloud subscription) that puts a threshold on the amount of resources to be invested in a query evaluation process. It is thus important to optimize the use of these resources.

 %We believe that the optimization of this process can occur at two levels: first at the level of agreed SLA exported by services.  Indeed, queries requesting the same services compositions will have clauses in their SLAs that are more conditions of use of the infrastructure (ie not storing the data produced by a service). Instead of recomputing the derived SLA every time, we propose to store it and reuse it for other queries.

%Second, pre-computed queries and partial results can be also stored in cache or in a persistent support. Rewriting results can be also stored and reduce execution and resources consumption time when evaluating similar queries. Storing or not such SLAs will depend on the cloud subscription of the user the issued the query (data access, intermediate storage capacity , cost of storage , etc ... ).

%As discussed in previous sections, the derived SLA associated to service compositions can have free measures that can only be evaluated at run-time. In order to do so, we assume that there is a monitoring system observing and aggregating events for computing resources consumption, execution and time cost, volumes of data transferred when services deliver results. These computations are used dynamically for instantiating free variables in the derived SLA and thus determine whether the contract is respected by the execution of a query. Since this is monitored dynamically, the evaluation of the query can be adapted if the SLA is not being respected anymore.

 %Furthermore, the way the integration will take place after the choice of the concrete service composition will require to enforce functional and non functional user and service preferences to take into account the integration constraints. In the example this has been illustrated by  the access control level constraints demanded by data services when they are implied in an integration process. Another important point is to consider the existing inter-cloud contracts as described in \cite{} to decide how to implement the integration it self.

>>>>

To answer the needs expressed above, we propose an SLA guided, security-aware continuous data provision and integration approach performed within a multi-cloud service oriented environment  (see
%starting from the processing of the query  to the delivery of the results.
Figure~\ref{fig:arch}). In this section, we present  step-by-step data integration process with the implied  SLA categories.

In our context, data are exposed as  a set of data services on different clouds. In our example of MOOC system presented above, several producers will supply equivalent content for a given period of time  that will be chosen with respect to the  preferences expressed by a consumer.

Given a query $Q_1$ expressed as Datalog-like program or an SQL-like expression, including spatio-temporal attributes and preferences.
For instance, the following query ``\textit{List of English poetry content providers that can provide commented Emily Dickinson poems that are close to my city and that are labelled as experts, where the total cost is not higher to 1 dollar, with using only services that preserve her anonymity}''.

The abstract Inter-Cloud Layer (ICL) is in charge of delivering a result to the received query. ICL will first identify a set of semantic functionalities that fit the  query and associated services, stored in the service description registry.

Assume that there are four content providers on English poetry that can be queried individually and two hubs that collect information from other sources like social network groups and hash tags.
Hubs will store content about given topics on  English poetry, available from particular producers (e.g., participants of a given course).
We represent these services by { e$_1$, \dots, e$_4$} and {h$_1$} and {h$_2$}. We suppose that hubs will have the same interface as individual service providers.
We also suppose the existence of two free location services exporting the following interface: {loc(IP) $\rightarrow$ $\langle$ X, Y$\rangle$}, meaning that given an IP address it returns a geographic position expressed as a pair of coordinates. 
All these services can potentially be combined for answering queries.

%shows the general architecture of an SLA guided data integration system. 
%It accesses data services which are data providers deployed in a cloud  that provide agreed SLA's. 
%The whole process is monitored to determine whether a computed SLA is being honoured while a query is evaluated.

@@

Each deployed service exposes an \textit{agreed SLA} that specifies the economic cost per call, the maximum number of calls that can be done per day, the availability of the service, the average response time when a method is called, the reliability, the privacy of the produced data (whether they can be stored or not), the precision, freshness and provenance of the produced data, as defined below:

@@

Some of these measures ({cost/call, maxCall/day}) are static and explicitly specified by the service provider. 
In contrast, the other measures should be computed by monitoring the conversations between the service and the applications that contact it.  It is the case for service reputation that will condition the way personal user data will be anonymized (Constraint1).  Some measures will be instantiated depending on the context of  the service invoked within a multi-cloud environment. It is the case for constraint 2 expressed above:   user data access credentials should be calculated according to the clouds hosting different data services.
An \textit{agreed SLA} is expressed through an  XML document using the specification WSLA (Web service level agreement \footnote{\footnotesize http://www.research.ibm.com/people/a/akeller/\-Data/WSLASpecV1-20030128.pdf}).

As aforementioned, an a-priori SLA enforcement mechanism is mandatory to assess data integration feasibility.
To this end, non-functional characteristics of the content provision services ({i.e e$_1$, ... e$_4$}) are calculated by specific cloud services (SCS) that enforce SLAs  and which should present an API, composed of a set of methods such as:

\begin{trivlist}\sf\footnotesize
\item[~-~S$_i$:]estim-cost$_i$(contents, req\_size, cost, prov\_size, loc), QoS preferences$_\mathit{user}$, agreedSLA$_i$.

@@

Given user expressed QoS preferences ({QoS preferences}$_\mathit{user}$) and {agreedSLA}$_i$, the first method allows to obtain for each service $e_i$ the estimation of the budget (in our example, of a given course (content)) and a required minimum size. In other words, it will returns the cost of the data, as well as its actual size and location.

 The second method is used to engage in a data service (in our example a course data service) and will be used later if the service is retained for a composition that answer the query.

%The user will ask for a given derived SLA to which the provider may agree.

ICL by asking SCS for each semantic functionality, obtain the set of potential services that satisfy either user preferences and  service offering described through their \textit{Agreed SLAs}. Indeed, the query $Q_1$ has two parts: the one that corresponds to the content asked for by the user  that will be solved by combining data services and the user preferences. For this query the user is ready to pay a maximum of \$1 as total cost; she requests that content providers be certified as experts (provenance) even if their data are not fresh; she requires a contract duration of one week. Besides, she wishes her privacy be saved by adapting the anonymity process with regard to service reputation. The maximum total cost will condition also which kind of privacy preserving algorithms to apply.

\begin{trivlist}\sf\footnotesize
\item[~-~QoS preferences$_\mathit{user}$: ] $\langle$cost $\leq$ \$1, availability, freshness = ``any'', provenance = ``certified'', location = ``close'', duration = 7 days, privacy-preserving=``reputation-based''$\rangle$. 
\end{trivlist}

Cloud providers also define their SLA contracts that specify the cost per request ({cost/request}), the volume of data that can be exchanged per month ({I/0 volume/month}), the cost of transferring data or applications within the same data center or between data centers ({datatransferCost/region}), the storage space ({storageSpace}). For example some cloud providers enable the customer to choose the zone to install PaaS services and deploy applications (e.g. zone 1 is Europe). If the customer wishes to deploy services in zone 1 but store data in zone 2 the transfer cost will change.

@@

Illustration inspired in the lowest contract provided by Azure\footnote{Azure is a trademark of Microsoft Corporation.}: 
 \begin{trivlist}\sf\footnotesize
\item[~-~cloudSLA:]  $\langle$0,05 cents per call, 8 GB I/0 volume/month, free, 1 GB storage$\rangle$. 
\end{trivlist}

@@

The next step supervised by the ICL consists in elaborating a set of potential composition specifications via a query rewriting process that complies both the user cloud SLAs and QoS preferences: cost, provenance, trust requirement, execution time, data confidentiality needs, \dots.
 This allows computing a result as follows: (i) User preferences (including cloud SLAs according to user subscriptions) and \textit{agreed SLAs} will be used to produce a \textit{derived SLA} to be associated to the query. These derived SLA will influence the choice of (data) services that will actually be used for building the result; (ii) Query rewriting for computing service compositions that can be used for  building query results. The \textit{agreed SLA} exposed by the chosen services is added as  conditions in  each re-written query expression; Two situations could arise: a full compliance of a composition with the \textit{derived SLA} or a partial compliance that requires a negotiation step. For both alternatives, an \textit{integration SLA} will be produced indicating the negotiation conditions and the implementation of the final integration; (iii) and Managing the integration process according to the \textit{integration SLA}.

%-----------------------------------------------------------------
\section{Derived SLA}
\label{sec:slaModel}
%-----------------------------------------------------------------

The key and original aspect of  our proposed data integration and provision process is  defined as a vertical mapping of user QoS preferences and agreed SLAs of potential com posable services. This  leads to a {\em derived SLA} that guides the evaluation of a query.

%A query has associated preferences  expressed as macroscopic constraints (i.e. user preferences statement): execution time, pay / no pay, data reliability, provenance, freshness, privacy-preserving constraints, partial/full results, delivery mode. These constraints are coupled with the profile of the user which is in general stated in her cloud subscription (amount of assigned storage space, number of requests, I/O transfered Mega bytes, etc.).

%As said before, we assume that services export their \textit{agreed SLAs} that define measures that can be either expressed as constants,  computed (dynamically) by monitoring the execution and conversations associated to services, and hybrid they can be statically stated  but they change at execution time.  A service  agreed SLA is expressed through an  XML document using the specification WSLA (Web service level agreement \footnote{\footnotesize http://www.research.ibm.com/people/a/akeller/\-Data/WSLASpecV1-20030128.pdf}). The service SLA measures  that we consider in our example are: cost, availability, freshness, provenance, location, duration of the engagement, privacy-preserving status. Other measures are associated to the conditions in which the service is called or to the precision and recall of their produced data given a request.

@@

%-----------------------------------------------------------------
\section{Towards an integration SLA}
\label{sec:queryProcessOpt}
%-----------------------------------------------------------------
 Our data provision and integration approach relies on data services deployed on one or different clouds providers and it is delivered as a distributed DaaS on top of the clouds involved.  This DaaS  uses resources from a cloud and this use  is  guided by an economic model (stated in a cloud subscription) that puts a threshold on the amount of resources to be invested in a query evaluation process. It is thus important to optimize the use of these resources.

The optimization of this process can occur at two levels: first at the level of agreed SLA exported by services.  Indeed, queries requesting the same services compositions will have clauses in their SLAs that are more conditions of use of the infrastructure (ie not storing the data produced by a service). Instead of recomputing the derived SLA every time, we propose to store it and reuse it for other queries. 
Second, pre-computed queries and partial results can be also stored in cache or in a persistent support. Rewriting results can be also stored and reduce execution and resources consumption time when evaluating similar queries. Storing or not such SLAs will depend on the cloud subscription of the user the issued the query (data access, intermediate storage capacity , cost of storage , etc ... ).

Furthermore, the way the integration will take place after the choice of the concrete service composition will require to enforce functional and non functional user and service preferences to take into account the integration constraints. In the example this has been illustrated by  the access control level constraints demanded by data services when they are implied in an integration process. Another important point is to consider the existing inter-cloud contracts as described in \cite{} to decide how to implement the integration it self. This leads to the build of the \textit{integration SLA}

%As discussed in previous sections, the derived SLA associated to service compositions can have free measures that can only be evaluated at run-time. In order to do so, we assume that there is a monitoring system observing and aggregating events for computing resources consumption, execution and time cost, volumes of data transferred when services deliver results. These computations are used dynamically for instantiating free variables in the derived SLA and thus determine whether the contract is respected by the execution of a query. Since this is monitored dynamically, the evaluation of the query can be adapted if the SLA is not being respected anymore.
%
